---
title: "Assignment_P1"
author: "Marcelladane"
date: "18 August 2017"
output:
  html_document: default
  pdf_document: default
---

##Statistical Inference Course Project

In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. You should
- Show the sample mean and compare it to the theoretical mean of the distribution.
- Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
- Show that the distribution is approximately normal.
In point 3, focus on the difference between the distribution of a large collection of random exponentials and the distribution of a large collection of averages of 40 exponentials.

###Simulation

Start by loading the necessary packages, here, just one for plot.
```{r}
packages <- c("ggplot2")
sapply(packages, require, character.only = TRUE, quietly = TRUE)

```

Now coming to the simulations, first let's set a seed (to create a set of values) and then set the needed constants (lambda, n and number of simulations, in this order)

```{r}
set.seed(123)
lambda <- 0.2
n <- 40
NS <- 1000
```

Now run the exponential distribution and plot it.
Calculate the means of distribution.
Plot a histogram with the mean values.

```{r plot1, echo=FALSE}
ExpD <- matrix(data=rexp(n*NS, lambda), nrow=NS)
ExpD_means <- data.frame(means=apply(ExpD, 1, mean))
qplot(ExpD_means$means, geom="histogram", xlab = "Means", main = "Distribution of averages of samples", fill=I("blue"), col=I("red"))
```

###Comparing Sample values versus Theoretical values

*MEANS*

The expected mean $\mu$ (mu) of a exponential distribution of rate $\lambda$ (lambda) is:

mu = 1 / lambda

And

Call X (in the classes was X line, but here I call X) the average of your 1000 simulations (NS) of 40 samples (n), so the theoretical mean is the mean of the means.

```{r}
mu = 1 / lambda
mu

MofMeans <- mean(ExpD_means$means)
MofMeans
```

They are not completely identical, but quite close means.

*VARIANCE*

The expected standard deviation $\sigma$ (sigma) of a exponential distribution of rate lambda is:

sigma = (1 / lambda)/ sqrt(n)

And

the variance (Var) of sigma is: Var=sigma^2
(to calculate the SD for the second example I will call sigma, SigmaX)

```{r}
sigma <- (1/lambda)/sqrt(n)
sigma

SigmaX <- sd(ExpD_means$means)
Var <- SigmaX^2
Var
```

Here the difference is a bit bigger then the mean one, but still fairly close.


###Finally, lets plot the distribution of means and see if it looks normal distribution

```{r plot2, echo=FALSE}
lambda_vals <- seq(4, 6, by=0.01)
coverage <- sapply(lambda_vals, function(lamb) {
    mu_hats <- rowMeans(matrix(rexp(n*NS, rate=0.2),
                               NS, n))
    ll <- mu_hats - qnorm(0.975) * sqrt(1/lambda**2/n)
    ul <- mu_hats + qnorm(0.975) * sqrt(1/lambda**2/n)
    mean(ll < lamb & ul > lamb)
})

library(ggplot2)
qplot(lambda_vals, coverage) + geom_hline(yintercept=0.95)
```

Here you can see that for selection of $\hat{\lambda}$ around 5, the average of the sample mean falls within the confidence interval at least 95% of the time. 

*Full code: https://github.com/marcelladane/statistical_inference*